---
title: "QBS_181_Midterm"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Hmisc)

data <- sasxport.get("DIQ_I.xpt")
```
\section{Q1}
(a) There is missing data in many of columns. For example, in the 'High Blood Sugar' column, there are
9498 out of 9575 missing values, in 'Craving for sweet/eating a lot of sugar' column, there are 9564 out of 9575 missing values. So we can't simply remove the rows with missing values. The summary of missing values are given below
```{r}
for (col in colnames(data)) {
  print(c('number of missing values: ',  sum(is.na(data[col]))))
  print(describe(data[col]))
}
```
From above summaries we can verify the counts of missing values are consistent with the description in the website.

(b) The values of some columns are categorical, meaning we can't simply replace the missing values by the mean. We noticed the check items, for example, the tag DIQ060U, DIQ229 and DIQ295 specifies from which column the remaining columns all have missing values. For example if ages less than 12 the columns from DIQ300S all have missing values. We further noticed that in column
'Feel could be at risk for diabetes' if the answers are 'No', all values up to column DIQ180 are missing. Therefore, we may only keep those answers 'Yes' in that column.

(c) The codes and missing values are less descriptive without referring to the website. We are going to replace the code with descriptive words as explained in the dataset. 



\section{Q2}
We first replace the missing values and some code with descriptive words:
```{r}
data$diq010[data$diq010 == 1] <- "Yes"
data$diq010[data$diq010 == 2] <- "No"
data$diq010[data$diq010 == 3] <- "Borderline"
data$diq010[data$diq010 == 7] <- "Refused"
data$diq010[data$diq010 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq010"]
```
```{r}
# do did040
data$did040[data$did040 == 666] <- "Less than 1 year"
data$did040[data$did040 == 777] <- "Refused"
data$did040[data$did040 == 999] <- "Don't know"
data[sample(nrow(data), 10), "did040"]
```

```{r}
# do diq160
data$diq160[data$diq160 == 1] <- "Yes"
data$diq160[data$diq160 == 2] <- "No"
data$diq160[data$diq160 == 7] <- "Refused"
data$diq160[data$diq160 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq160"]
```
`
```{r}
# do diq170
data$diq170[data$diq170 == 1] <- "Yes"
data$diq170[data$diq170 == 2] <- "No"
data$diq170[data$diq170 == 7] <- "Refused"
data$diq170[data$diq170 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq170"]
```

```{r}
# do diq172
data$diq172[data$diq172 == 1] <- "Yes"
data$diq172[data$diq172 == 2] <- "No"
data$diq172[data$diq172 == 7] <- "Refused"
data$diq172[data$diq172 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq172"]
```

```{r}
# do diq175a
data$diq175a[data$diq175a == 10] <- "Family history"
data$diq175a[data$diq175a == 77] <- "Refused"
data$diq175a[data$diq175a == 99] <- "Don't know"
data[sample(nrow(data), 10), "diq175a"]
```

```{r}
# do diq175b
data$diq175b[data$diq175b == 11] <- "Overweight"
data[sample(nrow(data), 10), "diq175b"]
```

```{r}
# do diq175c
data$diq175c[data$diq175c == 12] <- "Age"
data[sample(nrow(data), 10), "diq175c"]
```

```{r}
# do diq175d
data$diq175d[data$diq175d == 13] <- "Poor diet"
data[sample(nrow(data), 10), "diq175d"]
```

```{r}
# do diq175E
data$diq175e[data$diq175e == 14] <- "Race"
data[sample(nrow(data), 10), "diq175e"]
```

```{r}
# do diq175F
data$diq175f[data$diq175f == 15] <- "Had a baby weighed over 9 lbs. at birth"
data[sample(nrow(data), 10), "diq175f"]
```

```{r}
# do diq175g
data$diq175g[data$diq175g == 16] <- "Lack of physical activity"
data[sample(nrow(data), 10), "diq175g"]
```

```{r}
# do diq175H
data$diq175h[data$diq175h == 17] <- "High blood pressure"
data[sample(nrow(data), 10), "diq175h"]
```

```{r}
# do diq175I
data$diq175i[data$diq175i == 18] <- "High blood sugar"
data[sample(nrow(data), 10), "diq175i"]
```

```{r}
# do diq175j
data$diq175j[data$diq175j == 19] <- "High cholesterol"
data[sample(nrow(data), 10), "diq175j"]
```

```{r}
# do diq175K
data$diq175k[data$diq175k == 20] <- "Hypoglycemic"
data[sample(nrow(data), 10), "diq175k"]
```

```{r}
# do diq175L
data$diq175l[data$diq175l == 21] <- "Extreme hunger"
data[sample(nrow(data), 10), "diq175l"]
```

```{r}
# do diq175m
data$diq175m[data$diq175m == 22] <- "Tingling/numbness in hands or feet"
data[sample(nrow(data), 10), "diq175m"]
```

```{r}
# do diq175n
data$diq175n[data$diq175n == 23] <- "Blurred vision"
data[sample(nrow(data), 10), "diq175n"]
```

```{r}
# do diq175o
data$diq175o[data$diq175o == 24] <- "Increased fatigue"
data[sample(nrow(data), 10), "diq175o"]
```

```{r}
# do diq175p
data$diq175p[data$diq175p == 25] <- "Anyone could be at risk"
data[sample(nrow(data), 10), "diq175p"]
```

```{r}
# do diq175q
data$diq175q[data$diq175q == 26] <- "Doctor warning"
data[sample(nrow(data), 10), "diq175q"]
```

```{r}
# do diq175r
data$diq175r[data$diq175r == 27] <- "Other, specify"
data[sample(nrow(data), 10), "diq175r"]
```

```{r}
# do diq175s
data$diq175s[data$diq175s == 28] <- "Gestational diabetes"
data[sample(nrow(data), 10), "diq175s"]
```

```{r}
# do diq175t
data$diq175t[data$diq175t == 29] <- "Frequent urination"
data[sample(nrow(data), 10), "diq175t"]
```

```{r}
# do diq175u
data$diq175u[data$diq175u == 30] <- "Thirst"
data[sample(nrow(data), 10), "diq175u"]
```

```{r}
# do diq175v
data$diq175v[data$diq175v == 31] <- "Craving for sweet/eating a lot of sugar"
data[sample(nrow(data), 10), "diq175v"]
```

```{r}
# do diq175w
data$diq175w[data$diq175w == 32] <- "Medication"
data[sample(nrow(data), 10), "diq175w"]
```

```{r}
# do diq175x
data$diq175x[data$diq175x == 33] <- "Polycystic ovarian syndrome"
data[sample(nrow(data), 10), "diq175x"]
```

```{r}
# do diq180
data$diq180[data$diq180 == 1] <- "Yes"
data$diq180[data$diq180 == 2] <- "No"
data$diq180[data$diq180 == 7] <- "Refused"
data$diq180[data$diq180 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq180"]
```

```{r}
# do diq050
data$diq050[data$diq050 == 1] <- "Yes"
data$diq050[data$diq050 == 2] <- "No"
data$diq050[data$diq050 == 7] <- "Refused"
data$diq050[data$diq050 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq050"]
```

```{r}
# do diq060
data$diq060[data$diq060 == 666] <- "Less than 1 month"
data$diq060[data$diq060 == 777] <- "Refused"
data$diq060[data$diq060 == 999] <- "Don't know"
data[sample(nrow(data), 10), "diq060"]
```

```{r}
# do diq060U
data$diq060u[data$diq060u == 1] <- "Months"
data$diq060u[data$diq060u == 2] <- "Years"
data[sample(nrow(data), 10), "diq060u"]
```

```{r}
# do diq070
data$diq070[data$diq070 == 1] <- "Yes"
data$diq070[data$diq070 == 2] <- "No"
data$diq070[data$diq070 == 7] <- "Refused"
data$diq070[data$diq070 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq070"]
```

```{r}
# do diq230
data$diq230[data$diq230 == 1] <- "1 year ago or less"
data$diq230[data$diq230 == 2] <- "More than 1 year ago but no more than 2 years ago"
data$diq230[data$diq230 == 3] <- "More than 2 years ago but no more than 5 years ago"
data$diq230[data$diq230 == 4] <- "More than 5 years ago"
data$diq230[data$diq230 == 5] <- "Never"
data$diq230[data$diq230 == 7] <- "Refused"
data$diq230[data$diq230 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq230"]
```

```{r}
# do diq240
data$diq240[data$diq240 == 1] <- "Yes"
data$diq240[data$diq240 == 2] <- "No"
data$diq240[data$diq240 == 7] <- "Refused"
data$diq240[data$diq240 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq240"]
```


```{r}
# do diq260
data$diq260[data$diq260 == 0] <- "Never"
data$did040[data$diq260 == 777] <- "Refused"
data$diq260[data$diq260 == 999] <- "Don't know"
data[sample(nrow(data), 10), "diq260"]
```

```{r}
# do diq260u
data$diq260u[data$diq260u == 1] <- "Per day"
data$diq260u[data$diq260u == 2] <- "Per week"
data$diq260u[data$diq260u == 3] <- "Per month"
data$diq260u[data$diq260u == 4] <- "Per year"
data[sample(nrow(data), 10), "diq260u"]
```

```{r}
# do diq275
data$diq275[data$diq275 == 1] <- "Yes"
data$diq275[data$diq275 == 2] <- "No"
data$diq275[data$diq275 == 7] <- "Refused"
data$diq275[data$diq275 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq275"]
```

```{r}
# do diq280
data$diq280[data$diq280 == 777] <- "Refused"
data$diq280[data$diq280 == 999] <- "Don't know"
data[sample(nrow(data), 10), "diq280"]
```

```{r}
# do diq291
data$diq291[data$diq291 == 1] <- "Less than 6"
data$diq291[data$diq291 == 2] <- "Less than 7"
data$diq291[data$diq291 == 3] <- "Less than 8"
data$diq291[data$diq291 == 4] <- "Less than 9"
data$diq291[data$diq291 == 5] <- "Less than 10"
data$diq291[data$diq291 == 6] <- "Provider did not specify goal"
data$diq291[data$diq291 == 77] <- "Refused"
data$diq291[data$diq291 == 99] <- "Don't know"
data[sample(nrow(data), 10), "diq291"]
```

```{r}
# do diq300s
data$diq300s[data$diq300s == 7777] <- "Refused"
data$diq300s[data$diq300s == 9999] <- "Don't know"
data[sample(nrow(data), 10), "diq300s"]
```

```{r}
# do diq300d
data$diq300d[data$diq300d == 7777] <- "Refused"
data$diq300d[data$diq300d == 9999] <- "Don't know"
data[sample(nrow(data), 10), "diq300d"]
```


```{r}
# do did310d
data$did310d[data$did310d == 6666] <- "Provider did not specify goal"
data$did310d[data$did310d== 7777] <- "Refused"
data$did310d[data$did310d == 9999] <- "Don't know"
data[sample(nrow(data), 10), "did310d"]
```

```{r}
# do did320
data$did320[data$did320 == 5555] <- "Never heard of LDL"
data$did320[data$did320 == 6666] <- "Never had cholesterol test"
data$did320[data$did320 == 7777] <- "Refused"
data$did320[data$did320 == 9999] <- "Don't know"
data[sample(nrow(data), 10), "did320"]
```

```{r}
# do did330
data$did330[data$did330 == 6666] <- "Provider did not specify goal"
data$did330[data$did330 == 7777] <- "Refused"
data$did330[data$did330 == 9999] <- "Don't know"
data[sample(nrow(data), 10), "did330"]
```

```{r}
# do did341
data$did341[data$did341 == 0] <- "None"
data$did341[data$did341 == 7777] <- "Refused"
data$did341[data$did341 == 9999] <- "Don't know/not sure"
data[sample(nrow(data), 10), "did341"]
```

```{r}
# do did350
data$did350[data$did350 == 0] <- "None"
data$did350[data$did350 == 7777] <- "Refused"
data$did350[data$did350 == 9999] <- "Don't know"
data[sample(nrow(data), 10), "did350"]
```

```{r}
# do diq350u
data$diq350u[data$diq350u == 1] <- "Per day"
data$diq350u[data$diq350u == 2] <- "Per week"
data$diq350u[data$diq350u == 3] <- "Per month"
data$diq350u[data$diq350u == 4] <- "Per year"
data[sample(nrow(data), 10), "diq350u"]
```

```{r}
# do diq360
data$diq360[data$diq360 == 1] <- "Less than 1 month"
data$diq360[data$diq360 == 2] <- "1-12 months"
data$diq360[data$diq360 == 3] <- "13-24 months"
data$diq360[data$diq360 == 4] <- "Greater than 2 years"
data$diq360[data$diq360 == 5] <- "Never"
data$diq360[data$diq360 == 7] <- "Refused"
data$diq360[data$diq360 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq360"]
```

```{r}
# do diq080
data$diq080[data$diq080 == 1] <- "Yes"
data$diq080[data$diq080 == 2] <- "No"
data$diq080[data$diq080 == 7] <- "Refused"
data$diq080[data$diq080 == 9] <- "Don't know"
data[sample(nrow(data), 10), "diq080"]
```

Last, we will keep those answers 'Yes' in the 'Feel could be at risk for diabetes' column.
```{r}
small_data <- data[((data$diq010=="Borderline")|(data$diq160=="Yes")), ]

mydata <- small_data[which(small_data$diq172=="Yes"),]
```
After data cleaning, the size of data is down to 453 rows. We will store the data as the csv file and put it into database for further analysis.
```{r}
write.csv(mydata, 'DIQ_I_cleaned.csv')
```